Supplemental Material for: The evolution of gene expression in the term placenta of viviparous mammals

### 2016-11-17 manuscript/manuscript.Rnw
change title again
show the housekeeping genes which are core

\title{The evolution of the transcriptome of mammalian placentas}%
%
\author[affigb]{Don L. Armstrong\fnref{n1}}
\ead{don@donarmstrong.com}
\author[afflon]{Michael R. McGowen\fnref{n1}}
\ead{1mmcgowen1@gmail.com}
\author[affigb]{Amy Weckle}
\ead{aweckle@illinois.edu}
\author[affigb]{Priyadarshini Pantham}
\ead{pantham@illinois.edu}
\author[affwayne]{Jason Caravas}
\ead{jacaravas@gmail.com}
\author[affmich]{Dalen Agnew}
\ead{agnewd@dcpah.msu.edu}
\author[affuscd]{Kurt Benirschke}
\ead{kbenirschke@ucsd.edu}
\author[affapetrust]{Sue Savage-Rumbaugh}
\ead{suerumbaugh@gmail.com}
\author[affhaifa]{Eviatar Nevo}
\ead{nevo@research.haifa.ac.il}
\author[affnih,affwaynepath]{Chong J Kim}
\ead{cjkim@med.wayne.edu}
\author[affyale]{Günter P. Wagner}
\ead{gunter.wagner@yale.edu}
\author[affnih]{Roberto Romero}
\ead{prbchiefstaff@med.wayne.edu}
\author[affigb,affapetrust,affmip]{Derek E. Wildman\corref{cor1}}
\ead{wildmand@illinois.edu}
\cortext[cor1]{Corresponding author}
\address[affigb]{Institute for Genomic Biology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
\address[afflon]{School of Biological and Chemical Sciences, Queen Mary, University of London, London, UK}
\address[affwayne]{Center for Molecular Medicine and Genetics, Wayne State University School of Medicine, Detroit, MI, USA}
\address[affmich]{Diagnostic Center for Population and Animal Health, College of Veterinary Medicine, Michigan State University, East Lansing, MI, USA}
\address[affuscd]{Department of Pathology, University of California San Diego, La Jolla, CA, USA}
\address[affapetrust]{Great Ape Trust/Bonobo Hope Sanctuary, Des Moines, IA, USA}
\address[affnih]{Perinatology Research Branch, Eunice Kennedy Shriver National Institute of Child Health and Human Development (NICHD), National Institutes of Health (NIH), Detroit, MI, USA}
\address[affwaynepath]{Department of Pathology, Wayne State University School of Medicine, Detroit, MI, USA}
\address[affyale]{Department of Ecology and Evolutionary Biology and Yale Systems Biology Institute, Yale University, New Haven, CT, USA}
\address[affmip]{Department of Molecular and Integrative Physiology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
\address[affhaifa]{The Institute of Evolution, University of Haifa, Haifa, Israel}

@@

\begin{figure}[tbh]
<<core.placenta.transcriptome,fig.width=5,fig.height=8,out.height="0.8\\textheight",echo=FALSE>>=
cptl <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species),]
cptl <- cptl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]

>>>>

\title{The core transcriptome of mammalian placentas and the
  divergence of expression with placental shape}%
%
\author[affigb]{Don L. Armstrong\fnref{n1}}
\ead{don@donarmstrong.com}
\author[afflon]{Michael R. McGowen\fnref{n1}}
\ead{1mmcgowen1@gmail.com}
\author[affigb]{Amy Weckle}
\ead{aweckle@illinois.edu}
\author[affigb]{Priyadarshini Pantham}
\ead{pantham@illinois.edu}
\author[affwayne]{Jason Caravas}
\ead{jacaravas@gmail.com}
\author[affmich]{Dalen Agnew}
\ead{agnewd@dcpah.msu.edu}
\author[affuscd]{Kurt Benirschke}
\ead{kbenirschke@ucsd.edu}
\author[affapetrust]{Sue Savage-Rumbaugh}
\ead{suerumbaugh@gmail.com}
\author[affhaifa]{Eviatar Nevo}
\ead{nevo@research.haifa.ac.il}
\author[affnih,affwaynepath]{Chong J Kim}
\ead{cjkim@med.wayne.edu}
\author[affyale]{Günter P. Wagner}
\ead{gunter.wagner@yale.edu}
\author[affnih]{Roberto Romero}
\ead{prbchiefstaff@med.wayne.edu}
\author[affigb,affapetrust,affmip]{Derek E. Wildman\corref{cor1}}
\ead{wildmand@illinois.edu}
\cortext[cor1]{Corresponding author}
\address[affigb]{Institute for Genomic Biology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
\address[afflon]{School of Biological and Chemical Sciences, Queen Mary, University of London, London, UK}
\address[affwayne]{Center for Molecular Medicine and Genetics, Wayne State University School of Medicine, Detroit, MI, USA}
\address[affmich]{Diagnostic Center for Population and Animal Health, College of Veterinary Medicine, Michigan State University, East Lansing, MI, USA}
\address[affuscd]{Department of Pathology, University of California San Diego, La Jolla, CA, USA}
\address[affapetrust]{Great Ape Trust/Bonobo Hope Sanctuary, Des Moines, IA, USA}
\address[affnih]{Perinatology Research Branch, Eunice Kennedy Shriver National Institute of Child Health and Human Development (NICHD), National Institutes of Health (NIH), Detroit, MI, USA}
\address[affwaynepath]{Department of Pathology, Wayne State University School of Medicine, Detroit, MI, USA}
\address[affyale]{Department of Ecology and Evolutionary Biology and Yale Systems Biology Institute, Yale University, New Haven, CT, USA}
\address[affmip]{Department of Molecular and Integrative Physiology, University of Illinois at Urbana-Champaign, Urbana, IL, USA}
\address[affhaifa]{The Institute of Evolution, University of Haifa, Haifa, Israel}

@@

\begin{figure}[tbh]
<<core.placenta.transcriptome,fig.width=5,fig.height=8,out.height="0.8\\textheight",echo=FALSE>>=
cptl <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species) & housekeeping==FALSE,]
cptl <- cptl[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]

### 2016-11-17 manuscript/supplemental_material.Rnw
switch to acro from acronym
show the housekeeping genes which are core

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=15pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage{acronym}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature,citestyle=numeric-comp,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=blue,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\usepackage{xspace}
% \usepackage[below]{placeins}
% \makeatletter
% \AtBeginDocument{%
%   \expandafter\renewcommand\expandafter\subsection\expandafter
%     {\expandafter\@fb@secFB\subsection}%
%   \newcommand\@fb@secFB{\FloatBarrier
%     \gdef\@fb@afterHHook{\@fb@topbarrier \gdef\@fb@afterHHook{}}}%
%   \g@addto@macro\@afterheading{\@fb@afterHHook}%
%   \gdef\@fb@afterHHook{}%
% }
% \makeatother
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

>>>>

\documentclass[english,12pt]{article}
\usepackage{fontspec}
\setmainfont{FreeSerif}
\setsansfont{FreeSans}
\setmonofont{FreeMono}
\usepackage[letterpaper,left=1in,right=1in,top=1in,bottom=1in,headheight=15pt]{geometry}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[bf]{caption}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{float}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{setspace}
\usepackage{wrapfig}
% \usepackage{txfonts}
\usepackage{acro}
\usepackage{array}
\usepackage{dcolumn}
\usepackage{adjustbox}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage[backend=biber,natbib=true,hyperref=true,style=nature,citestyle=numeric-comp,autocite=inline]{biblatex}
\addbibresource{references.bib}
\usepackage[noblocks,auth-sc]{authblk}
\usepackage[hyperfigures,bookmarks,colorlinks,linkcolor=black,urlcolor=blue,citecolor=blue]{hyperref}
%\usepackage[hyperfigures,bookmarks,colorlinks,citecolor=black,filecolor=black,linkcolor=black,urlcolor=black]{hyperref}
% Floats at end for journals
% \usepackage[nolists,figuresfirst,nomarkers,heads,nofighead,notabhead]{endfloat}
% \AtBeginFigures{\section*{Figure Legends}\renewcommand{\efloatseparator}{\hbox{}}}
% \AtBeginTables{\section*{Tables}\renewcommand{\efloatseparator}{\mbox{}}}
\usepackage{float}
\floatstyle{plaintop}
\restylefloat{table}
\usepackage[nomargin,inline,draft]{fixme}
\newcommand{\DLA}[1]{\textcolor{red}{\fxnote{DLA: #1}}}
\usepackage[x11names,svgnames]{xcolor}
\usepackage{texshade}
\usepackage{tikz}
\usepackage{nameref}
\usepackage{zref-xr,zref-user}
\usepackage{cleveref}
\usepackage{xspace}
% \usepackage[below]{placeins}
% \makeatletter
% \AtBeginDocument{%
%   \expandafter\renewcommand\expandafter\subsection\expandafter
%     {\expandafter\@fb@secFB\subsection}%
%   \newcommand\@fb@secFB{\FloatBarrier
%     \gdef\@fb@afterHHook{\@fb@topbarrier \gdef\@fb@afterHHook{}}}%
%   \g@addto@macro\@afterheading{\@fb@afterHHook}%
%   \gdef\@fb@afterHHook{}%
% }
% \makeatother
\def\newblock{\hskip}
\newenvironment{paperquote}{%
  \begin{quote}%
     \it
  }%
  {\end{quote}}
\renewcommand{\textfraction}{0.15}
\renewcommand{\topfraction}{0.85}
\renewcommand{\bottomfraction}{0.65}
\renewcommand{\floatpagefraction}{0.60}
%\renewcommand{\baselinestretch}{1.8}
\newenvironment{enumerate*}%
  {\begin{enumerate}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{enumerate}}
\newenvironment{itemize*}%
  {\begin{itemize}%
    \setlength{\itemsep}{0pt}%
    \setlength{\parskip}{0pt}}%
  {\end{itemize}}
\oddsidemargin 0.0in 
\textwidth 6.5in
\raggedbottom
\clubpenalty = 10000
\widowpenalty = 10000
% make the afill smaller
\renewcommand\Affilfont{\itshape\small}
\renewcommand\cite[1]{\autocite{#1}}
\pagestyle{fancy}

@@

\begin{figure}[tbh]
<<core.placenta.transcriptome.housekeeping,fig.width=5,fig.height=10,out.height="0.8\\textheight",echo=FALSE>>=
cptlh <-
    core.placenta.transcriptome.long[!grepl("(min|median|max|all)_expression|all_but_",
                                            Species) & housekeeping==TRUE,]
cptlh <- cptlh[,Species:=factor(gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",capfirst(Species)),
                              levels=gsub("^([^ ])[^ ]* +([^ ]*)","\\1. \\2",
                                          capfirst(names(species.names))))]

print(ggplot(cptlh,aes(x=human_name,y=FPKM))+
      geom_boxplot(outlier.shape=NA)+
      scale_y_log10(labels=trans_format("log10", math_format(10^.x)))+
      geom_point(aes(color=Species,shape=Species))+
      scale_shape_manual("Species",
                         guide=guide_legend(#nrow=1,
                                            keywidth=unit(3,"mm"),
                                            keyheight=unit(3,"mm"),
                                            label.theme=element_text(face="italic",
                                                                     angle=0,
                                                                     size=9),
                                            title.position="top"
                                            ),
                         values=0:18)+
      # scale_color_discrete(NULL)+
      coord_flip()+
      xlab("Human Ortholog Name")+
      ylab("Gene Expression (FPKM)")+
      theme_bw() +
      theme(legend.position=c(0.7,0.15),
            axis.text.y=element_text(face="italic",size=5),
            legend.margin=unit(0,"mm"),
            legend.justification="left",
            plot.margin=unit(c(0,0,0,0),"mm")
            ))
@ 
\caption{Housekeeping genes excluded from the Core Placenta Transcriptome.
  \Sexpr{cptlh[,length(unique(human_name))]}
  1:1 \human ortholog genes expressed ($\mathrm{FPKM} \ge 10$) in the
  placenta of all \Sexpr{star.logs[,length(unique(species))]} species
  studied which are housekeeping genes~%
  \cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  ordered (from top to bottom) by median expression.}
\label{fig:core_placenta_transcriptome_housekeeping}
\end{figure}
